---
title: "Human activity prediction"
author: "Nataliya Skrypnyuk"
date: "22/11/2015"
output: html_document
---
### Executive summary

The purpose of this report is to present a strategy for determining, which movement with a barbell has been made based on the measurements from a wearable device. There are 5 movements classes and over 100 parameters are being measured.

We first clean the data, by taking out all measurements with lot's of unavailable data, and then apply the random forest methods to the part of the training set. The resulting prediction is very good both on the part of the training set used for training and on the left out part. When we apply our method to the testing set, then all 20 test cases are predicted correctly.

### Data cleaning

```{r, echo=TRUE, eval=FALSE}
training<-read.csv(file="pml-training.csv",header=TRUE)
testing<-read.csv(file="pml-testing.csv",header=TRUE)

#inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]

a<-c()
for (i in colnames(training)) {
  a<-c(a,sum(is.na(training[,i]))/length(training[,i]))
}

a

for (i in colnames(training)) {
  a<-c(a,sum(training[,i]=="")/length(training[,i]))
  }

a

training<-training[,a==0]
testing<-testing[,a==0]


library(caret)
modfit <- train(classe ~ ., data = training, trControl=trainControl(method = "cv", number = 3), method="rf")
output<-pre
a<-c()
for (i in colnames(training)) {
  a<-c(a,sum(training[,i]=="")/length(training[,i]))
  }
training<-training[,a==0]
testing<-testing[,a==0]
training<-training[,-c(1:7)]
testing<-testing[,-c(1:7)]

inTrain = createDataPartition(training$classe, p = 0.7)[[1]]
datatrain<-training[inTrain,]
datatest<-training[-inTrain,]

rf<-randomForest(classe~.,data=datatrain)
rfpredict<-predict(rf,datatest)
sum(rfpredict==datatest$classe)/length(datatest$classe)

intcolnames<-intersect(colnames(testing2),colnames(training))
newtesting<-testing2[,intcolnames]
rfpredict<-predict(rf,newtesting)

#B  A  B  A  A  E  D  B  A  A  B  C  B  A  E  E  A  B  B  B 

```

### Model fitting

```{r, echo=TRUE, eval=FALSE}
training<-read.csv(file="pml-training.csv",header=TRUE)
testing<-read.csv(file="pml-testing.csv",header=TRUE)

#inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]

a<-c()
for (i in colnames(training)) {
  a<-c(a,sum(is.na(training[,i]))/length(training[,i]))
  }
training<-training[,a==0]
testing<-testing[,a==0]
library(caret)
modfit <- train(classe ~ ., data = training, trControl=trainControl(method = "cv", number = 3), method="rf")
output<-pre
a<-c()
for (i in colnames(training)) {
  a<-c(a,sum(training[,i]=="")/length(training[,i]))
  }
training<-training[,a==0]
testing<-testing[,a==0]
training<-training[,-c(1:7)]
testing<-testing[,-c(1:7)]

inTrain = createDataPartition(training$classe, p = 0.7)[[1]]
datatrain<-training[inTrain,]
datatest<-training[-inTrain,]

rf<-randomForest(classe~.,data=datatrain)
rfpredict<-predict(rf,datatest)
sum(rfpredict==datatest$classe)/length(datatest$classe)

intcolnames<-intersect(colnames(testing2),colnames(training))
newtesting<-testing2[,intcolnames]
rfpredict<-predict(rf,newtesting)

#B  A  B  A  A  E  D  B  A  A  B  C  B  A  E  E  A  B  B  B 

```

### Model validation

```{r, echo=TRUE, eval=FALSE}
training<-read.csv(file="pml-training.csv",header=TRUE)
testing<-read.csv(file="pml-testing.csv",header=TRUE)

#inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]

a<-c()
for (i in colnames(training)) {
  a<-c(a,sum(is.na(training[,i]))/length(training[,i]))
  }
training<-training[,a==0]
testing<-testing[,a==0]
library(caret)
modfit <- train(classe ~ ., data = training, trControl=trainControl(method = "cv", number = 3), method="rf")
output<-pre
a<-c()
for (i in colnames(training)) {
  a<-c(a,sum(training[,i]=="")/length(training[,i]))
  }
training<-training[,a==0]
testing<-testing[,a==0]
training<-training[,-c(1:7)]
testing<-testing[,-c(1:7)]

inTrain = createDataPartition(training$classe, p = 0.7)[[1]]
datatrain<-training[inTrain,]
datatest<-training[-inTrain,]

rf<-randomForest(classe~.,data=datatrain)
rfpredict<-predict(rf,datatest)
sum(rfpredict==datatest$classe)/length(datatest$classe)

intcolnames<-intersect(colnames(testing2),colnames(training))
newtesting<-testing2[,intcolnames]
rfpredict<-predict(rf,newtesting)

#B  A  B  A  A  E  D  B  A  A  B  C  B  A  E  E  A  B  B  B 

```
